\textsf{Are large language models superhuman chemists?}

### 2024-10-24 src/tex/appendix.tex
feat: plot chembench corpus and tiny subset

\Cref{fig:all_questions_models_completely_correct_radar_overall} shows the performance of the models on the different topics of the \chembench corpus.
The general pattern of performance varies significantly between the different topics and is also observed when the models are evaluated on the entire corpus.
However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on \chembenchmini.

@@

\begin{table}
    \caption{\textbf{Performance of the models on \chembenchmini.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
    \resizebox{\textwidth}{!}{
    \variable{output/performance_table_human_subset.tex}
    }
    \label{tab:performance_table_human_subset}
\end{table}

@@

\Cref{fig:performance_per_topic_tiny} shows the same analysis on \chembenchmini.

\begin{figure}[htb]
    \centering
    \includegraphics{figures/performance_per_topic_tiny.pdf}
    \caption{\textbf{Fraction of  correctly answered questions per data source on \chembenchmini.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources. The performance is measured as the fraction of questions answered completely correctly by the models. A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none were answered completely correctly.
        We see that the performance of the models varies significantly between the different data sources. For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans. However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
    }
    \label{fig:performance_per_topic_tiny}
    \script{analyze_performance_per_source.py}
\end{figure}

@@

In the applications, human participants were presented with molecules as rendered drawings and SMILES strings. \LaTeX\xspace equations and chemical equations were rendered using MathJax (\Cref{fig:screenshots}).

>>>>

\Cref{fig:all_questions_models_completely_correct_radar_overall} shows the performance of the models on the different topics of the \chembench corpus.
The general pattern of performance varies significantly between the different topics and is also observed when the models are evaluated on the entire corpus.
However, since some subjects are composed of questions from different sources, the ranking of the models is, in some instances, different from the one on the \enquote{tiny} subset.

@@

\begin{table}
    \caption{\textbf{Performance of the models on the \enquote{tiny} subset of the \chembench corpus.} The table shows the fraction of questions answered completely correctly by the models for different skills and difficulty levels.}
    \resizebox{\textwidth}{!}{
    \variable{output/performance_table_human_subset.tex}
    }
    \label{tab:performance_table_human_subset}
\end{table}

@@

\Cref{fig:performance_per_topic_tiny} shows the same analysis on the \enquote{tiny} subset.

\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \enquote{tiny} subset. The relative performance difference between both is actually quite similar across most models. This makes \enquote{tiny} subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.

\begin{figure}[htb]
    \centering
    \includegraphics{figures/performance_per_topic_tiny.pdf}
    \caption{\textbf{Fraction of completely correctly answered questions per data source on the \enquote{tiny} subset.} The heatmap shows, in color, the fraction of questions answered completely correctly by different systems for some of our data sources. The performance is measured as the fraction of questions answered completely correctly by the models. A score of one (red) indicates that all questions were answered completely correctly, while a score of zero (blue) indicates that none were answered completely correctly.
        We see that the performance of the models varies significantly between the different data sources. For instance, it is interesting to observe that questions sourced based on textbooks seem easier for the leading models than for humans. However, this performance does not correlate with performance on other sources, e.g., semi-programmatically created tasks such as questions about the number of signals in an \gls{nmr} spectrum.
    }
    \label{fig:performance_per_topic_tiny}
    \script{analyze_performance_per_source.py}
\end{figure}

\begin{figure}[htb]
    \centering
    \includegraphics{figures/corpus_human_comparison.pdf}
    \caption{\textbf{
        Performance of the models on the \chembench corpus and \enquote{tiny} subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both \chembench corpus and the \enquote{tiny} subset.
        We see that the model ranking remains fairly consistent across both sets}  
    }
    \label{fig:performance_corpus_and_tiny}
    \script{corpus_humanset_performance.py}
\end{figure}

@@

In the applications, human participants were presented with molecules as rendered drawings and SMILES strings. \LaTeX\xspace equations and chemical equations were rendered using MathJax (\Cref{fig:screenshots}.
